SPA: An Apache Spark Pilot scheduler for HPC applications

### 2019-02-27 paper.tex

\section{Introduction}
\section{Materials and Methods}\label{sec:methods}
    \todo{walltime parameter of workflow}
    \subsection{Added value of pilot scheduling}
        \todo{execution of a batch cluster vs dynamic (pilot) cluster}
    \subsection{Robust masters}
        \todo{scala standalone vs our pyspark workaround. kill masters in experiments}
    \subsection{Checkpointing}
        \todo{metric for determining how often to checkpoint based on cluster size}
    \subsection{Job arrays}
        \todo{need to kill idle workers. may not want all workers to be running at once.}
    \subsection{Example application}
        \todo{incrementation with varying task durations}
\section{Discussion}\label{sec:discussion}
\section{Conclusion}\label{sec:conclusion}

>>>>

\section{Introduction}

In the recent years, neuroimaging Big Data has become more accessible, as a
result of the growing number of open-data initiatives\cite{openneuro, hcp, ukbiobank}. However, the
processing of such large datasets using standard neuroimaging pipelines has 
become unfeasable. As a result, many researchers are moving towards Big Data 
engines, such as Apache Spark and Dask\cite{spark, dask}, to create their processing pipelines. 
Unlike traditional neuroimaging engines, big data engines were designed to be 
executed on dedicated commodity infrastructures using Big Data schedulers. Although
researchers may have access to clouds or local workstations, they still largely 
rely on High Performance Computing (HPC) infrastructure to process their data.

High-Performance Computing infrastructure differs from the commodity infrastructure 
typically used by Big Data platforms in that it is not dedicated. That is, there 
are multiple users running a variety of different applications and all using the 
same resources. Schedulers used on HPC infrastructure include PBS, SGE, Slurm, HTCondor and 
TORQUE\cite{schedulers}. The majority of these being batch submission schedulers
in which the number of resources are requested by the user and the specified 
program is executed on the infrastructure.

Another popular HPC scheduling approach,
as implemented by HTCondor, is known as pilot-scheduling. Rather than requesting
all necessary resources in a single batch call, a pilot-scheduling system will request
multiple instances of subsets of the required resources. Processing starts occuring 
as soon as a minimal amount of resources are allocated, and can be replenished if 
resources are lost. Such a scheduling approach
decreases the time a user spends in the queue waiting for resources as less resources
are being requested per instance.\todo{need citation}. Many scientific workflow
engines are compatible with pilot-scheduling schedulers \cite{nipype and others} and
some others exclusively use pilot-scheduling approaches \cite{Pegasus and PSOM}.

Apache Spark is a popular Big Data framework, commonly used in both industrial
and academic settings. Although it is a Scala-based framework, it also has APIs
for Java, Python (PySpark) and R. What made it powerful over its predecessors is their main
abstraction known as the Resilient Distributed Dataset (RDD), which enabled
in-memory processing of pipelines by scheduling the tasks to where the data is 
located. Through the use of RDDs, it also became possible to execute iterative 
workflows -- something not easily doable in older frameworks such as Hadoop MapReduce.
Schedulers for Spark include its built-in standalone schedule, Yet Another Resource Negotiator (YARN),
and Mesos \cite{yarn, mesos}. Out of the three schedulers, only Mesos can be used
as an HPC scheduler. However, Mesos is not frequently used as a scheduler in HPC
environments.

Although still not widely adopted possibly due to the difficulties in adopting Big Data frameworks on HPC \todo{better explain the difficulties...setting walltimes etc. also explain neuroimaging applications},
projects, such as Thunder, a PySpark-compatible library of tools from processing
images and time-series data of various fields, including neuroscience, exist. In addition,
Spark has been used in various research projects to process neuroimaging data \cite{Boubela, ariel's paper, maybe the simulation one}.

The recommended method for launching Spark applications on our HPC scheduler 
involves batch requesting all the necessary resources and launching a Standalone
Spark cluster once the resources have been requested. This could significantly increase
the scheduling time of Spark-based applications as a large amount of resources 
may be necessary to process Big Data. Rather than requesting all the resources 
at once, a Spark cluster is launched with a subset of the resources and is expanded as 
more resources get allocated. This, in turn, may reduce the overall processing of
an application as less resources are requested at once and are therefore more likely
to be scheduled faster.

There have been some efforts to use pilot-scheduling approaches with Spark applications
running on HPC \cite{jha and spark on pbs paper}. However, these projects are either
not open-source and not necessarily compatible with Slurm.

Here we present SPA, a pilot-scheduling library for launching Spark applications 
on HPC infrastracture. This libary currently enables the launching of Spark applications
on Slurm-based clusters by dividing the total amount of resources required and requesting
multiple instances of subsets of the resources at a time. When a single instance is started,
a Standalone Spark cluster is started. Built-in fault-tolerance of Spark masters 
and workers is leveraged, in addition to Driver fault tolerance for Scala-based applications.
When nodes are lost due to walltime expiration, additional pilot nodes are launched until
the driver completes. Using this system, it is more feasible to run large neuroimaging Spark applications on
HPC systems.

\section{Materials and Methods}\label{sec:methods}
    \todo{walltime parameter of workflow}
    \subsection{Added value of pilot scheduling}
        \todo{execution of a batch cluster vs dynamic (pilot) cluster}
    \subsection{Robust masters}
        \todo{scala standalone vs our pyspark workaround. kill masters in experiments}
    \subsection{Checkpointing}
        \todo{metric for determining how often to checkpoint based on cluster size}
    \subsection{Job arrays}
        \todo{need to kill idle workers. may not want all workers to be running at once.}
    \subsection{Example application}
        \todo{incrementation with varying task durations}
\section{Discussion}\label{sec:discussion}
\section{Conclusion}\label{sec:conclusion}
